---
title: "NDI_GLSS7_DendrogramScore"
author: "Zachary Gersten, PhD, MPH"
date: "January 2024"
output: 
---

# NDI_GLSS7_DendrogramScore: Creating the functional diversity measurement in R Vegan

__Required Input Files__
  - *nutrienttraitmatrix.csv* - Nutrient trait matrix
  - *hhfoodexp.csv* - Household food expenditure 

__Information__
This script ...

  1) Reads in the GLss 7 nutrient trait matrix and selects nutrient traits
  2) Calculates the Euclidean distance for each food pair and stores values in a square distance matrix
  3) Applies hierarchical Clustering using UPGMA to the distance matrix and creates the food dendrogram
  4) Assesses the total NDI GLSS 7 score
  5) Assesses the raw household NDI GLSS 7 score

__Output__
  - distancematrix.csv
  - fooddendrogram.jpeg
  
```{r Load packages, include= FALSE, warning=FALSE, message=FALSE}
library(vegan)
library(readxl)
library(tidyverse)
library(writexl)
library(dendextend)
library(circlize)
library(ggplot2)
library(ggdendro)
```  
  
  
```{r Set working directory} 
setwd("C:/Users/zachary.gersten/Box/USDA/2_USDA-ARS/Indicator Development/GLSS 7 NDI/GitHub")
```   


```{r Load Data, include=FALSE}
nutrienttraitmatrix <- read_csv("nutrienttraitmatrix.csv")
nutrienttraitmatrixframe <- as.data.frame(nutrienttraitmatrix)
rownames(nutrienttraitmatrixframe) <- nutrienttraitmatrixframe$foodname
nutrienttraitmatrixframe <- nutrienttraitmatrixframe[, -1]
```


```{r Create distance matrix}
distancematrix <- vegdist(nutrienttraitmatrixframe, method="euclidean")
write.csv(distancematrix, "distancematrix.csv")
```


```{r Hierarchal cluster of distance matrix}  
foodclusters <- hclust(distancematrix, method = "average")
jpeg(file = 'fooddendrogram.jpeg', quality = 100, width = 2000, height = 1000)
plot(foodclusters, hang = -1, cex = 1)
```


```{r Display food dendrogram}
ggdendrogram(foodclusters, rotate = TRUE, theme_dendro = FALSE)

```

```{r Find the sum of the lengths of the connecting segments in a dendrogram as the total NDI GLSS 7 score}
treeheight(foodclusters)
```


```{r Load Data, include=FALSE}
hhfoodexp <- read_csv("hhfoodexp.csv")
hhfoodexpframe <- as.data.frame(hhfoodexp)
rownames(hhfoodexpframe) <- hhfoodexpframe$hid
hhfoodexpframe <- hhfoodexpframe[, -1]
```


```{r Find the tree height for each household for the raw household NDI GLss7 score}
rawhhNDIscore <- treedive(hhfoodexpframe, foodclusters)
write.csv(rawhhNDIscore, "rawhhNDIscore.csv")
```